\addcontentsline{toc}{chapter}{Acknowledgments}
\thispagestyle{empty}

\vspace*{20mm}

\begin{center}
    {\usekomafont{sectioning}\usekomafont{section} Acknowledgments}
\end{center}

\vspace{10mm}

%TODO: Acknowledgments
\noindent
First and foremost, I would like to thank God for the guidance, strength and perseverance that made it possible to complete this work.
\newline

\noindent
I would also like to express my sincere gratitude to all the people who supported, inspired, and challenged me throughout the work on this thesis.
\newline

\noindent
I would like to thank my supervisor Mihail Stoian for the continuous guidance, insightful feedback, and trust in my ideas. The discussions we had, ranging from high-level algorithmic ideas to low-level implementation details, were invaluable and significantly shaped the direction and quality of this work. I am especially grateful for the encouragement to explore optimizations deeply and to question existing assumptions about compression performance.

\noindent
I am deeply thankful for the practical insights, constructive criticism, and technical expertise, particularly regarding algorithmic correctness and performance evaluation. The feedback I received helped turn initial ideas into a well-structured and measurable contribution.
\newline

\noindent
I am grateful to the broader research and open-source community whose work laid the foundation for this thesis. The availability of high-quality implementations, documentation, and discussions around compression techniques made it possible to build upon existing work and focus on meaningful improvements.

\noindent
A special thanks goes to everyone who shared tools, datasets, or prior experience that helped accelerate experimentation and evaluation.
\newline

\noindent
Finally, I want to thank my friends and family for their constant support, patience, and encouragement throughout this journey. Their motivation helped me stay focused and enthusiastic, especially during long debugging sessions and performance tuning phases.
\newline

\noindent
This thesis would not have been possible without the collective contributions of all these people, and I am truly thankful for their support and inspiration.

\cleardoublepage{}
